Multi-product Batch Processing Time Maximization Problem

### 2023-08-22 manuscript.tex
updating paper
explaining better exact method

@@

After relaxation of the problem, it is possible to consider the factory stock and the outlets stock as single stock, so we have:

\begin{equation}
E_i = O_i + I_i
\end{equation}

where $E_i$ is the sum of the quantity stored at the factory and the quantity sent to the outlets of the product $i$.

@@

\emph{Solution for the example presented before}

Applying the method for solving the example presented in section \ref{sec:MBPTMP}, we have:

$T' \leq \lfloor{1000 \textrm{ g} / (60 \textrm{ g}/\textrm{min})}\rfloor$

$T' \leq 16 \textrm{ min}$ \quad for A

$T' \leq \lfloor{500 \textrm{ g} / (40 \textrm{ g}/\textrm{min})}\rfloor$

$T' \leq 12 \textrm{ min}$ \quad for B

so:

$T' = 12 \textrm{ min}$

Then:

$S_A = 1000 \textrm{ g} - 12 \textrm{ min} * 60 \textrm{ g}/\textrm{min} = 280 \textrm{ g}$

$S_B = 500 \textrm{ g} - 12 \textrm{ min} * 40 \textrm{ g}/\textrm{min} = 20 \textrm{ g}$

And so we have:

$T'' \leq \lfloor{(3000 \textrm{ g} + 600 \textrm{ g} + 280 \textrm{ g}) / (60 \textrm{ g}/\textrm{min})}\rfloor$

$T'' \leq 64 \textrm{ min}$ \quad for A

$T'' \leq \lfloor{(2000 \textrm{ g} + 600 \textrm{ g} + 20 \textrm{ g}) / (40 \textrm{ g}/\textrm{min})}\rfloor$

$T'' \leq 65 \textrm{ min}$ \quad for B

$T'' \leq \lfloor{(1000 \textrm{ g} + 3000 \textrm{ g} + 300 \textrm{ g}) / (100 \textrm{ g}/\textrm{min})}\rfloor$

$T'' \leq 43 \textrm{ min}$ \quad for A and B

$T'' = 43 \textrm{ min}$

So we have:

$T = 55 \textrm{ min}$

As $T<\textrm{Z}$, then $T = 55 \textrm{ min}$ will certainly be the optimal solution.

In this case we will have:

$P_A = 3300 \textrm{ g}$ e $P_B=2200 \textrm{ g}$

$D_A = 1000 \textrm{ g}$ e $D_B = 500 \textrm{ g}$

$E_A = 2300 \textrm{ g}$ e $E_B = 1700 \textrm{ g}$

@@

To test the developed model and analytical solution method, we created a solver in C++ named COPSolver (V01\_20230809, lib multiproduct-batch-processing-time-maximization-problem) and a solver in LINGO named MPBPTM.lng both available at https://github.com/tbfraga/COPSolver. The tests were performed on a notebook with an Intel i7 processor. We tested the solvers developed for the benchmarks presented on Tables \ref{tab:MBPTMP001}, \ref{tab:MBPTMP002} and \ref{tab:MBPTMP003}, and for randon benchmarks generated by the following functions: \\

@@

\begin{table}[h]
\begin{center}
\begin{footnotesize}
\begin{tabular}[c]{l r r r r}
problem & LINGO solver & time (s) & analytical method & time (s) \\
\cline {1-5} \\
MBPTM 2 & 55 & 0.04 & 55 & 0.001 \\
MBPTM 3 & 48 & 0.06 & 48 & 0.001 \\
MBPTM 10 & 30 & 0.09 & 30 & 0.001 \\
RMBPTM 20 & 100 & 0.10 & 100 & 0.002 \\
RMBPTM 50 & 98 & 0.07 & 98 & 0.002\\
RMBPTM 100 & 98 & 0.11 & 98 & 0.002\\
RMBPTM 1,000 & 78 & 9.59 & 78 & 0.013\\
RMBPTM 2,000 & 70 & 3.24 & 70 & 0.027\\
RMBPTM 5,000 & 70 & 15.28 & 70 & 0.079 \\
RMBPTM 10,000 & 70 & 58.22 & 70 & 0.185 \\
\cline {1-5} \\
\end{tabular}
\caption{Results obtained with the LINGO solver and the analytical method}
\label{tab:results}
\end{footnotesize}
\end{center}
\end{table}

>>>>

O método proposto neste artigo para solução do MBPTM problem é composto por duas partes distintas, que são: a) cálculo do tempo máximo de processamento do lote multi-produtos; e b) cálculo e distribuição da produção entre a demanda, os outlets e o estoque em fábrica. A seguir, explicamos cada uma destas partes detalhadamente.

\noindent \emph{cálculo do tempo máximo de processamento do lote multi-produtos:}

It is possible to split the batch processing time into two time slots:

\noindent where $T'$ é o tempo máximo de processamento cuja produção será utilizada exclusivamente para atendimento da demanda.

Neste caso, sendo $D_{i}^{'}$ a quantidade produzida do produto $i$ após $T'$ unidades de tempo de processamento do lote, temos que:

\begin{equation}
\label{eq:s02}
D_{i}^{'} - \textrm{p}_i * T'  = 0 \quad \forall i
\end{equation}

\begin{equation}
\label{eq:s03}
D_{i}^{'} \in Z \quad \forall i
\end{equation}

\noindent e, como $D_{i}^{'} \leq D_i$, $\forall i$, temos que:

\begin{equation}
\label{eq:s04}
D_{i}^{'} \leq \textrm{UD}_i\quad \forall i
\end{equation}

Das equações em (\ref{eq:s02}), (\ref{eq:s03}) e (\ref{eq:s04}), temos:

\begin{equation}
\label{eq:s05}
T' \leq \lfloor{\textrm{UD}_i / \textrm{p}_i}\rfloor \quad \forall i
\end{equation}

E, como desejamos máximo valor para $T'$, de (\ref{eq:s05}), temos:

\begin{equation}
\label{eq:s06}
T' = \min_i\{\lfloor{\textrm{UD}_i / \textrm{p}_i}\rfloor\}
\end{equation}

Observer que poderá haver uma sobra correspondente à demanda não atendida pela produção ocorrida na primeira parte do tempo de processamento do lote, $T'$. Esta sobra, $S_i$, será definida de acordo com as equações em (\ref{eq:s07}):

\begin{equation}
\label{eq:s07}
S_i = \textrm{UD}_i - D_{i}^{'} \quad \forall i
\end{equation}

É possível também considerar que os outlets são uma extensão do estoque da fábrica. Portanto it is possible to consider the factory stock of product $i$ and the outlets stock $i$ as single stock for the product $i$, de forma que:

\begin{equation}
\label{eq:s08}
E_i = O_i + I_i
\end{equation}

\begin{equation}
\label{eq:s09}
E_i \leq \textrm{UO}_i + \textrm{UI}_i
\end{equation}

\begin{equation}
\label{eq:s10}
\sum_i {E_i} \leq \textrm{O} + \textrm{I}
\end{equation}

\begin{equation}
\label{eq:s11}
E_i \in Z \quad \forall i
\end{equation}

\noindent where $E_i$ is the sum of factory stock and outlets stock of the product $i$.

Como o tempo $T'$ foi utilizado apenas para produção que irá atender à demanda, $T''$ será o restante do tempo de processamento, ou seja, o tempo máximo de processamento utilizado para produção que irá atender à sobra da demanda, será enviada para os outlets e estocada em fábrica. Assim temos que:

\begin{equation}
\label{eq:s12}
E_i + S_i - \textrm{p}_i * T''  = 0 \quad \forall i
\end{equation}

Portanto, das restrições em (\ref{eq:s09}) e equações em (\ref{eq:s12}), temos:

\begin{equation}
\label{eq:s13}
T''  \leq (\textrm{UO}_i + \textrm{UI}_i + S_i) / \textrm{p}_i \quad \forall i
\end{equation}

\noindent e, das restrições em (\ref{eq:s10}) e equações em (\ref{eq:s12}), temos:

\begin{equation}
\label{eq:s14}
 T'' \leq (\textrm{O} + \textrm{I} + \sum_i S_i) / \sum_i \textrm{p}_i
\end{equation}

Finalmente, como $T''$ é um valor inteiro e queremos maximizar esse valor, então:

\begin{equation}
\label{eq:s15}
 T'' = \lfloor\min\{ \min_i\{ (\textrm{UO}_i + \textrm{UI}_i + S_i) / \textrm{p}_i\}, (\textrm{O} + \textrm{I} + \sum_i S_i) / \sum_i \textrm{p}_i \}\rfloor
\end{equation}

\noindent Assim, $T$ pode ser calculado conforme equação em (\ref{eq:s17}):

\begin{equation}
\label{eq:s17}
 T = \lfloor \min_i\{{\textrm{UD}_i / \textrm{p}_i}\} + \min\{ \min_i\{ (\textrm{UO}_i + \textrm{UI}_i + S_i) / \textrm{p}_i\}, (\textrm{O} + \textrm{I} + \sum_i S_i) / \sum_i \textrm{p}_i \}\rfloor
\end{equation}

@@

After relaxation of the problem,

@@

@@

To test the developed model and analytical solution method, we created a solver in C++ named COPSolver (V01\_20230814, lib multiproduct-batch-processing-time-maximization-problem) and a solver in LINGO named MBPTM.lng both available at https://github.com/tbfraga/COPSolver. The tests were performed on a notebook with an Intel i7 processor. We tested the solvers developed for the benchmarks presented on Tables \ref{tab:MBPTMP001}, \ref{tab:MBPTMP002} and \ref{tab:MBPTMP003}, and for randon benchmarks generated by the following functions: \\

@@

\begin{table}[h]
\begin{center}
\begin{footnotesize}
\begin{tabular}[c]{l r r r r}
problem & LINGO solver & time (s) & analytical method & time (s) \\
\cline {1-5} \\
MBPTM 2 & 55 & 0.03 & 55 & $<$ 0.01 \\
MBPTM 3 & 48 & 0.03 & 48 & $<$ 0.01 \\
MBPTM 10 & 30 & 0.05 & 30 & $<$ 0.01 \\
RMBPTM 20 & 100 & 0.06 & 100 & $<$ 0.01 \\
RMBPTM 50 & 98 & 0.08 & 98 & $<$ 0.01 \\
RMBPTM 100 & 98 & 0.11 & 98 & $<$ 0.01 \\
RMBPTM 1,000 & 78 & 1.20 & 78 & $<$ 0.01 \\
RMBPTM 2,000 & 70 & 3.25 & 70 & $<$ 0.01 \\
RMBPTM 5,000 & 70 & 15.12 & 70 & $<$ 0.02 \\
RMBPTM 10,000 & 70 & 55.19 & 70 & $<$ 0.04 \\
\cline {1-5} \\
\end{tabular}
\caption{Results obtained with the LINGO solver and the analytical method}
\label{tab:results}
\end{footnotesize}
\end{center}
\end{table}
